---
title: "Study Strategy Analyses"
author: "Rachel Rineman, Amanda Mae Woodward"
date: "9/29/2022"
output: html_document
---

```{r library}
library(readr)
library(expss)
library(tidyverse)
library(lme4)
library(viridis)
```

```{r}
StudyDat  <-read.csv("~/Desktop/Study Strategies Fixed.csv")
colnames(StudyDat)[1]<- "SubjectNumber"
colnames(StudyDat)[2]<- "ClassYear"
colnames(StudyDat)[4]<- "RaceEthnicity"
```

# Question 1: Do students in this class use effective study strategies? 
```{r priorSplit} 
PriorStudyDat<- separate(StudyDat,MC1_PriorStrategyDunlosky, into= c("PS1", "PS2", "PS3", "PS4", "PS5", "PS6"), sep= ",")
```

```{r switchtoLong }
PriorStudyLong<- pivot_longer(PriorStudyDat, cols= c("PS1", "PS2", "PS3", "PS4", "PS5", "PS6"), names_to= "orderMention", values_to= "Strategies")

PriorStudyLong$Strategies<- as.factor(PriorStudyLong$Strategies)
levels(PriorStudyLong$Strategies)
levels(PriorStudyLong$Strategies)<- c("Distributed", "Highlight", "Imagery", "Interleaved","Mnemonic","Mnemonic", "Reread", "SelfExplain", "Summary", "TestPractice","Distributed", "Highlight","Interleaved", "Mnemonic", "Reread","SelfExplain", "Summary", "TestPractice",  "TestPractice")

```

```{r Priorstudydatacleaning}
PriorStudyLong$MC1_PriorStrategyUtility<- ifelse(PriorStudyLong$Strategies == "Distributed", "High", PriorStudyLong$MC1_PriorStrategyUtility)
PriorStudyLong$MC1_PriorStrategyUtility<- ifelse(PriorStudyLong$Strategies == "TestPractice", "High", PriorStudyLong$MC1_PriorStrategyUtility)
PriorStudyLong$MC1_PriorStrategyUtility<- ifelse(PriorStudyLong$Strategies == "Interleaved", "High", PriorStudyLong$MC1_PriorStrategyUtility)
PriorStudyLong$MC1_PriorStrategyUtility<- ifelse(PriorStudyLong$Strategies == "Interrogation", "Moderate", PriorStudyLong$MC1_PriorStrategyUtility)
PriorStudyLong$MC1_PriorStrategyUtility<- ifelse(PriorStudyLong$Strategies == "SelfExplain", "Moderate", PriorStudyLong$MC1_PriorStrategyUtility)
PriorStudyLong$MC1_PriorStrategyUtility<- ifelse(PriorStudyLong$Strategies == "Reread", "Low", PriorStudyLong$MC1_PriorStrategyUtility)
PriorStudyLong$MC1_PriorStrategyUtility<- ifelse(PriorStudyLong$Strategies == "Highlight", "Low", PriorStudyLong$MC1_PriorStrategyUtility)
PriorStudyLong$MC1_PriorStrategyUtility<- ifelse(PriorStudyLong$Strategies == "Summary", "Low", PriorStudyLong$MC1_PriorStrategyUtility)
PriorStudyLong$MC1_PriorStrategyUtility<- ifelse(PriorStudyLong$Strategies == "Mnemonic", "Low", PriorStudyLong$MC1_PriorStrategyUtility)
PriorStudyLong$MC1_PriorStrategyUtility<- ifelse(PriorStudyLong$Strategies == "Imagery", "Low", PriorStudyLong$MC1_PriorStrategyUtility)
```

## Question 1: Do students in this class use effective study strategies?

### Do students use effective strategies more often than chance at the beginning of the course? 

#### Hypothesis: We hypothesize that students will not report using effective strategies at the beginning of the course

#### Descriptive Statistics About Study Strategies 
```{r descriptivePriorStrat}
cro(PriorStudyLong$Strategies)
library(viridis)
PriorStudyLong$MC1_PriorStrategyUtility<- as.factor(PriorStudyLong$MC1_PriorStrategyUtility)
levels(PriorStudyLong$MC1_PriorStrategyUtility)

class(PriorStudyLong$MC1_PriorStrategyUtility)
PriorStudyLong$MC1_PriorStrategyUtility<-ordered(PriorStudyLong$MC1_PriorStrategyUtility, levels= c("Low", "Moderate", "High"))
PsGraph<- cbind.data.frame(PriorStudyLong$Strategies, PriorStudyLong$MC1_PriorStrategyUtility)
PsGraph<- na.omit(PsGraph)
colnames(PsGraph)<- c("Strategies", "MC1_PriorStrategyUtility")
ggplot(PsGraph, aes(Strategies, fill=MC1_PriorStrategyUtility))+ geom_bar() + theme_classic()+ scale_fill_viridis(option="D", discrete=TRUE)+facet_wrap(~MC1_PriorStrategyUtility)+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

#### Descriptive Statistics About Study Strategy Utilities  
```{r descriptivePriorStratEffective}
cro(PriorStudyLong$MC1_PriorStrategyUtility)
ggplot(PsGraph, aes(MC1_PriorStrategyUtility))+geom_bar()+ theme_classic()
```
#of the 150 student responses (some students provided more than 1), 74 mentioned high utility 64 mentioned low, 10 mentioned moderate

#### Analysis: Generalized Linear Model Logistic Regression (Preliminary)
Dependent Variable: Utility (1= high utility; 0 = not high utility)
Independent Variable: Demographic Variables
```{r}
PriorStudyLong$Utility<- ifelse(PriorStudyLong$MC1_PriorStrategyUtility== "High", 1, 0)
# need to make column to indicate if student used *any* high utility strategy
colnames(PriorStudyDat)
PriorStudyDat<- as.data.frame(PriorStudyDat)
for(i in 5:10){
  PriorStudyDat[,i]<- as.factor(PriorStudyDat[,i])
}
class(PriorStudyDat[,5])
levels(PriorStudyDat[,5])[5]<- "Reread"
levels(PriorStudyDat[,6])<- c("Distributed","Highlight","Reread","SelfExplain","Summary","TestPractice")
levels(PriorStudyDat[,7])<- c("Distributed","Highlight","Imagery","Interleaved","Mnemonic","Reread","SelfExplain","Summary","TestPractice")
levels(PriorStudyDat[,8])<- c("Distributed","Highlight","Mnemonic","Reread","SelfExplain")
levels(PriorStudyDat[,9])<-c("Distributed","SelfExplain","TestPractice")
levels(PriorStudyDat[,10])<- c("SelfExplain","TestPractice")
PriorStudyDat$OverallUtility<-NA
```

```{r}
PriorStudyDat$PS1<-ifelse(is.na(PriorStudyDat$PS1)==TRUE, "none", levels(PriorStudyDat$PS1)[PriorStudyDat$PS1])
PriorStudyDat$PS2<-ifelse(is.na(PriorStudyDat$PS2)==TRUE, "none", levels(PriorStudyDat$PS2)[PriorStudyDat$PS2])
PriorStudyDat$PS3<-ifelse(is.na(PriorStudyDat$PS3)==TRUE, "none", levels(PriorStudyDat$PS3)[PriorStudyDat$PS3])
PriorStudyDat$PS4<-ifelse(is.na(PriorStudyDat$PS4)==TRUE, "none", levels(PriorStudyDat$PS4)[PriorStudyDat$PS4])
PriorStudyDat$PS5<-ifelse(is.na(PriorStudyDat$PS5)==TRUE, "none", levels(PriorStudyDat$PS5)[PriorStudyDat$PS5])
PriorStudyDat$PS6<-ifelse(is.na(PriorStudyDat$PS6)==TRUE, "none", levels(PriorStudyDat$PS6)[PriorStudyDat$PS6])

for (j in 5:10){
  for(i in 1:nrow(PriorStudyDat)){
    if(PriorStudyDat[i,j]== "TestPractice"){
      PriorStudyDat$OverallUtility[i]<-1
    }
    if(PriorStudyDat[i,j]=="Distributed"){
      PriorStudyDat$OverallUtility[i]<-1
    }
    if(PriorStudyDat[i,j]=="Interleaved"){
       PriorStudyDat$OverallUtility[i]<-1
    }
    else{print("whoops!")}
  }
}

PriorStudyDat$OverallUtility<- ifelse(is.na(PriorStudyDat$OverallUtility)==TRUE,0, PriorStudyDat$OverallUtility)
PriorStudyDat$ClassYear<- as.factor(PriorStudyDat$ClassYear)
levels(PriorStudyDat$ClassYear)
levels(PriorStudyDat$ClassYear)<- c("Do Not Wish to Disclose", "First Year Student/Freshman", "First Year Student/Freshman", "Fourth Year Student/Senior", "Fourth Year Student/Senior", "Second Year Student/Sophomore", "Second Year Student/Sophomore", "Third Year Student/Junior", "Third Year Student/Junior")

PriorStudyDat$Gender<-as.factor(PriorStudyDat$Gender)
levels(PriorStudyDat$Gender)<- c("Another gender identity not listed here", "Cisgender Woman", "Cisgender Man", "Cisgender Woman", "Do Not Wish to Disclose", "Do Not Wish to Disclose", "Gender Nonbinary", "Transgender Woman")
levels(PriorStudyDat$Gender)

PriorStudyDat$RaceEthnicity<- as.factor(PriorStudyDat$RaceEthnicity)
levels(PriorStudyDat$RaceEthnicity)<- c("Asian", "Black or African American", "Black or African American", "Do Not Wish to Disclose", "Hispanic, Latino/Latina/Latinx or other Spanish Origin", "Middle Eastern or North African", "White or European American", "White or European American" )
levels(PriorStudyDat$RaceEthnicity)

glm1<-glm(OverallUtility~ClassYear+RaceEthnicity+Gender, data=PriorStudyDat)
summary(glm1)
glm2<-glm(OverallUtility~ClassYear, data=PriorStudyDat)
summary(glm2)

```

### Analysis: Binomial Test 
```{r}
summary(as.factor(PriorStudyDat$OverallUtility))
binom.test(51, (51+15))
```
**Note:** Students are (not?) using highly effective strategies more often than chance before the course

```{r explorePrior}
PriorStudyDat$lowImpact<- NA

for (j in 5:10){
  for(i in 1:nrow(PriorStudyDat)){
    if(PriorStudyDat[i,j]== "Highlight"){
      PriorStudyDat$lowImpact[i]<-1
    }
    if(PriorStudyDat[i,j]=="Imagery"){
      PriorStudyDat$lowImpact[i]<-1
    }
    if(PriorStudyDat[i,j]=="Mnemonic"){
       PriorStudyDat$lowImpact[i]<-1
    }
     if(PriorStudyDat[i,j]=="Reread"){
       PriorStudyDat$lowImpact[i]<-1
     }
     if(PriorStudyDat[i,j]=="Summary"){
       PriorStudyDat$lowImpact[i]<-1
    }
    else{print("whoops!")}
  }
}

PriorStudyDat$lowImpact<- ifelse(is.na(PriorStudyDat$lowImpact)== TRUE, "none",PriorStudyDat$lowImpact)


summary(as.factor(PriorStudyDat$lowImpact))

binom.test(44,66)

# students are statistically choosing low utility

PriorStudyDat$JustHighImpact<- ifelse(PriorStudyDat$OverallUtility==1 & PriorStudyDat$lowImpact !=1, 1, 0)
summary(as.factor(PriorStudyDat$JustHighImpact))
binom.test(13,66)

PriorStudyDat$JustLowImpact<- ifelse(PriorStudyDat$OverallUtility!=1 & PriorStudyDat$lowImpact==1, 1, 0)
summary(as.factor(PriorStudyDat$JustLowImpact))
binom.test(6,66)
```

#overall take away: more students than chance use high impact (and low impact strategies) at start of semester. Students seem to use a mix of both- not rely on one or the other 


Question 2: Will students use more effective strategies at the end of the course compared to at the beginning? 
We hypothesize that students will use more effective strategies as the semester progresses, compared to the beginning of the semester. 
### Data Clean 

#### MC 2
```{r}
colnames(StudyDat)
MC2StudyDat<- separate(StudyDat,MC2_UsedStrategyDunlosky, into= c("PS1", "PS2", "PS3", "PS4", "PS5", "PS6"), sep= ", ")

levels(MC2StudyDat$PS1)
MC2StudyDat$PS1<-as.factor(MC2StudyDat$PS1)
levels(MC2StudyDat$PS1)[6]<- "Reread"

MC2StudyDat$PS2<-as.factor(MC2StudyDat$PS2)
levels(MC2StudyDat$PS2)
levels(MC2StudyDat$PS2)<- c("Summary", "Distributed", "Highlight", "Highlight","Imagery","Interleaved","Interrogation","Mnemonic","Reread","Reread","SelfExplain", "Summary", "Summary", "TestPractice")
MC2StudyDat$PS3<-as.factor(MC2StudyDat$PS3)
levels(MC2StudyDat$PS3)<- c("NA", "Highlight", "Imagery", "Mnemonic", "Reread", "SelfExplain","Summary", "Summary", "TestPractice"  )
levels(MC2StudyDat$PS3)

MC2StudyDat$PS4<-as.factor(MC2StudyDat$PS4)
levels(MC2StudyDat$PS4)
levels(MC2StudyDat$PS4)[2]<-"Reread" 
MC2StudyDat$PS5<-as.factor(MC2StudyDat$PS5)
levels(MC2StudyDat$PS5)
MC2StudyDat$PS6<-as.factor(MC2StudyDat$PS6)
levels(MC2StudyDat$PS6)
```

```{r}
MC2_CurrentLongDat<- pivot_longer(MC2StudyDat,cols= c("PS1", "PS2", "PS3", "PS4", "PS5"), names_to= "orderMention", values_to= "Strategies")
```

#Longdat helps with graphs
```{r Priorstudydatacleaning}
colnames(MC2_CurrentLongDat)
MC2_CurrentLongDat$MC2_UsedUtility<- ifelse(MC2_CurrentLongDat$Strategies == "Distributed", "High", MC2_CurrentLongDat$MC2_UsedUtility)
MC2_CurrentLongDat$MC2_UsedUtility<- ifelse(MC2_CurrentLongDat$Strategies == "TestPractice", "High", MC2_CurrentLongDat$MC2_UsedUtility)
MC2_CurrentLongDat$MC2_UsedUtility<- ifelse(MC2_CurrentLongDat$Strategies == "Interleaved", "High", MC2_CurrentLongDat$MC2_UsedUtility)
MC2_CurrentLongDat$MC2_UsedUtility<- ifelse(MC2_CurrentLongDat$Strategies == "Interrogation", "Moderate", MC2_CurrentLongDat$MC2_UsedUtility)
MC2_CurrentLongDat$MC2_UsedUtility<- ifelse(MC2_CurrentLongDat$Strategies == "SelfExplain", "Moderate", MC2_CurrentLongDat$MC2_UsedUtility)
MC2_CurrentLongDat$MC2_UsedUtility<- ifelse(MC2_CurrentLongDat$Strategies == "Reread", "Low", MC2_CurrentLongDat$MC2_UsedUtility)
MC2_CurrentLongDat$MC2_UsedUtility<- ifelse(MC2_CurrentLongDat$Strategies == "Highlight", "Low", MC2_CurrentLongDat$MC2_UsedUtility)
MC2_CurrentLongDat$MC2_UsedUtility<- ifelse(MC2_CurrentLongDat$Strategies == "Summary", "Low", MC2_CurrentLongDat$MC2_UsedUtility)
MC2_CurrentLongDat$MC2_UsedUtility<- ifelse(MC2_CurrentLongDat$Strategies == "Mnemonic", "Low", MC2_CurrentLongDat$MC2_UsedUtility)
MC2_CurrentLongDat$MC2_UsedUtility<- ifelse(MC2_CurrentLongDat$Strategies == "Imagery", "Low", MC2_CurrentLongDat$MC2_UsedUtility)
cbind.data.frame(MC2_CurrentLongDat$Strategies, MC2_CurrentLongDat$MC2_UsedUtility)
```


#### Descriptive Statistics About Study Strategies MC2
```{r descriptivePriorStrat}
colnames(MC2_CurrentLongDat)
cro(MC2_CurrentLongDat$MC2_UsedUtility)
MC2_CurrentLongDat$MC2_UsedUtility<- as.factor(MC2_CurrentLongDat$MC2_UsedUtility)
levels(MC2_CurrentLongDat$MC2_UsedUtility)

class(MC2_CurrentLongDat$MC2_UsedUtility)
MC2_CurrentLongDat$MC2_UsedUtility<-ordered(MC2_CurrentLongDat$MC2_UsedUtility, levels= c("Low", "Moderate", "High"))
UsGraph<- cbind.data.frame(MC2_CurrentLongDat$Strategies, MC2_CurrentLongDat$MC2_UsedUtility)
UsGraph<- na.omit(UsGraph)
colnames(UsGraph)<- c("Strategies", "MC2_UsedUtility")
ggplot(UsGraph, aes(Strategies, fill=MC2_UsedUtility))+ geom_bar() + theme_classic()+ scale_fill_viridis(option="D", discrete=TRUE)+facet_wrap(~MC2_UsedUtility)+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

```{r}
cro(MC2_CurrentLongDat$MC2_UsedUtility)
class(MC2_CurrentLongDat$MC2_UsedUtility)
ggplot(UsGraph, aes(MC2_UsedUtility))+geom_bar(color= "skyblue")+ theme_classic()
```
#Of the 104 student reponses (some students provided more than one), 50 mentioned high utility, 16 mentioned moderate utility, and 55 mentioned low utility. 

#### Analysis: Generalized Linear Model Logistic Regression (Preliminary)
Dependent Variable: Utility (1= high utility; 0 = not high utility)
Independent Variable: Demographic Variables
```{r}
MC2_CurrentLongDat$OverallUtility<- ifelse(MC2_CurrentLongDat$MC2_UsedUtility== "High", 1, 0)
colnames(MC2_CurrentLongDat)

MC2StudyDat$PS1<-ifelse(is.na(MC2StudyDat$PS1)==TRUE, "none", levels(MC2StudyDat$PS1)[MC2StudyDat$PS1])
MC2StudyDat$PS2<-ifelse(is.na(MC2StudyDat$PS2)==TRUE, "none", levels(MC2StudyDat$PS2)[MC2StudyDat$PS2])
MC2StudyDat$PS3<-ifelse(is.na(MC2StudyDat$PS3)==TRUE, "none", levels(MC2StudyDat$PS3)[MC2StudyDat$PS3])
MC2StudyDat$PS4<-ifelse(is.na(MC2StudyDat$PS4)==TRUE, "none", levels(MC2StudyDat$PS4)[MC2StudyDat$PS4])
MC2StudyDat$PS5<-ifelse(is.na(MC2StudyDat$PS5)==TRUE, "none", levels(MC2StudyDat$PS5)[MC2StudyDat$PS5])
MC2StudyDat$PS6<-ifelse(is.na(MC2StudyDat$PS6)==TRUE, "none", levels(MC2StudyDat$PS6)[MC2StudyDat$PS6])

MC2StudyDat$OverallUtility<-NA
colnames(MC2StudyDat)

for (j in 16:20){
  for(i in 1:nrow(MC2StudyDat)){
    if(MC2StudyDat[i,j]== "TestPractice"){
      MC2StudyDat$OverallUtility[i]<-1
    }
    if(MC2StudyDat[i,j]=="Distributed"){
      MC2StudyDat$OverallUtility[i]<-1
    }
    if(MC2StudyDat[i,j]=="Interleaved"){
       MC2StudyDat$OverallUtility[i]<-1
    }
    else{print("whoops!")}
  }
}

MC2StudyDat$OverallUtility<- ifelse(is.na(MC2StudyDat$OverallUtility)==TRUE,0, MC2StudyDat$OverallUtility)


MC2StudyDat$ClassYear<- as.factor(MC2StudyDat$ClassYear)
levels(MC2StudyDat$ClassYear)
levels(MC2StudyDat$ClassYear)<- c("Do Not Wish to Disclose", "First Year Student/Freshman", "First Year Student/Freshman", "Fourth Year Student/Senior", "Fourth Year Student/Senior", "Second Year Student/Sophomore", "Second Year Student/Sophomore", "Third Year Student/Junior", "Third Year Student/Junior")

MC2StudyDat$Gender<-as.factor(MC2StudyDat$Gender)
levels(MC2StudyDat$Gender)<- c("Another gender identity not listed here", "Cisgender Woman", "Cisgender Man", "Cisgender Woman", "Do Not Wish to Disclose", "Do Not Wish to Disclose", "Gender Nonbinary", "Transgender Woman")
levels(MC2StudyDat$Gender)

MC2StudyDat$RaceEthnicity<- as.factor(MC2StudyDat$RaceEthnicity)
levels(MC2StudyDat$RaceEthnicity)<- c("Asian", "Black or African American", "Black or African American", "Do Not Wish to Disclose", "Hispanic, Latino/Latina/Latinx or other Spanish Origin", "Middle Eastern or North African", "White or European American", "White or European American" )

levels(MC2StudyDat$RaceEthnicity)

glm3<-glm(OverallUtility~ClassYear+RaceEthnicity+Gender, data=MC2StudyDat)
summary(glm3)
glm4<-glm(OverallUtility~ClassYear, data=MC2StudyDat)
summary(glm4)
```

### Analysis: Binomial Test 
```{r}
summary(as.factor(MC2StudyDat$OverallUtility))
binom.test(38, (38+28))
```
**Note**: Students are not using effective strategies more often than chance at time point 2. 

```{r explorePrior}
MC2StudyDat$lowImpact<- NA
colnames(MC2StudyDat)

for (j in 16:20){
  for(i in 1:nrow(MC2StudyDat)){
    if(MC2StudyDat[i,j]== "Highlight"){
      MC2StudyDat$lowImpact[i]<-1
    }
    if(MC2StudyDat[i,j]=="Imagery"){
      MC2StudyDat$lowImpact[i]<-1
    }
    if(MC2StudyDat[i,j]=="Mnemonic"){
       MC2StudyDat$lowImpact[i]<-1
    }
     if(MC2StudyDat[i,j]=="Reread"){
       MC2StudyDat$lowImpact[i]<-1
     }
     if(MC2StudyDat[i,j]=="Summary"){
       MC2StudyDat$lowImpact[i]<-1
    }
    else{print("whoops!")}
  }
}

MC2StudyDat$lowImpact<- ifelse(is.na(MC2StudyDat$lowImpact)== TRUE, 0, MC2StudyDat$lowImpact)


summary(as.factor(MC2StudyDat$lowImpact))

binom.test(36,66)

#students are not statistically choosing low utility over high utility at time point two.

MC2StudyDat$JustHighImpact<- ifelse(MC2StudyDat$OverallUtility==1 & MC2StudyDat$lowImpact !=1, 1, 0)
summary(as.factor(MC2StudyDat$JustHighImpact))

binom.test(46,66)

#students are statistically choosing high utility strategies at time point two. 

MC2StudyDat$JustLowImpact<- ifelse(MC2StudyDat$OverallUtility!=1 & MC2StudyDat$lowImpact==1, 1, 0)
summary(as.factor(MC2StudyDat$JustLowImpact))
binom.test(18,66)
 #students are statistically choosing low utility strategies at time point two.
```
#Takeaway: At time point two, students continue to choose both high and low strategies, like they did at time point one. 


### MC 3 
```{r}
StudyDat$MC3_UsedDunloskyStrategy
MC3StudyDat<- separate(StudyDat,MC3_UsedDunloskyStrategy, into= c("PS1", "PS2", "PS3", "PS4", "PS5"), sep= ", ")

MC3StudyDat$PS1<- as.factor(MC3StudyDat$PS1)
levels(MC3StudyDat$PS1)<-c("Distributed", "Highlight", "Imagery", "Interleaved", "Interleaved", "Interrogation", "NA", "Reread", "SelfExplain", "Summary", "Summary", "TestPractice", "TestPractice")
MC3StudyDat$PS2<- as.factor(MC3StudyDat$PS2)
levels(MC3StudyDat$PS2)
levels(MC3StudyDat$PS2)<- c("Distributed", "Highlight", "Highlight", "Imagery", "Interrogation", "Mnemonic", "Reread", "SelfExplain", "Summary", "TestPractice", "TestPractice")
MC3StudyDat$PS3<- as.factor(MC3StudyDat$PS3)
levels(MC3StudyDat$PS3)[3]<-"Reread"
levels(MC3StudyDat$PS3)
MC3StudyDat$PS4<- as.factor(MC3StudyDat$PS4)
levels(MC3StudyDat$PS4)
MC3StudyDat$PS5<- as.factor(MC3StudyDat$PS5)
levels(MC3StudyDat$PS5)
```

```{r}
MC3_CurrentLongDat<- pivot_longer(MC3StudyDat,cols= c("PS1", "PS2", "PS3", "PS4", "PS5"), names_to= "orderMention", values_to= "Strategies")

levels(MC3_CurrentLongDat$Strategies)

colnames(MC3_CurrentLongDat)

MC3_CurrentLongDat$MC3_UsedUtility<- ifelse(MC3_CurrentLongDat$Strategies == "Distributed", "High", MC3_CurrentLongDat$MC3_UsedUtility)
MC3_CurrentLongDat$MC3_UsedUtility<- ifelse(MC3_CurrentLongDat$Strategies == "TestPractice", "High", MC3_CurrentLongDat$MC3_UsedUtility)
MC3_CurrentLongDat$MC3_UsedUtility<- ifelse(MC3_CurrentLongDat$Strategies == "Interleaved", "High", MC3_CurrentLongDat$MC3_UsedUtility)
MC3_CurrentLongDat$MC3_UsedUtility<- ifelse(MC3_CurrentLongDat$Strategies == "Interrogation", "Moderate", MC3_CurrentLongDat$MC3_UsedUtility)
MC3_CurrentLongDat$MC3_UsedUtility<- ifelse(MC3_CurrentLongDat$Strategies == "SelfExplain", "Moderate", MC3_CurrentLongDat$MC3_UsedUtility)
MC3_CurrentLongDat$MC3_UsedUtility<- ifelse(MC3_CurrentLongDat$Strategies == "Reread", "Low", MC3_CurrentLongDat$MC3_UsedUtility)
MC3_CurrentLongDat$MC3_UsedUtility<- ifelse(MC3_CurrentLongDat$Strategies == "Highlight", "Low", MC3_CurrentLongDat$MC3_UsedUtility)
MC3_CurrentLongDat$MC3_UsedUtility<- ifelse(MC3_CurrentLongDat$Strategies == "Summary", "Low", MC3_CurrentLongDat$MC3_UsedUtility)
MC3_CurrentLongDat$MC3_UsedUtility<- ifelse(MC3_CurrentLongDat$Strategies == "Mnemonic", "Low", MC3_CurrentLongDat$MC3_UsedUtility)
MC3_CurrentLongDat$MC3_UsedUtility<- ifelse(MC3_CurrentLongDat$Strategies == "Imagery", "Low", MC3_CurrentLongDat$MC3_UsedUtility)
```
```

```{r}
cro(MC3_CurrentLongDat$MC3_UsedUtility)
MC3_CurrentLongDat$MC3_UsedUtility<- as.factor(MC3_CurrentLongDat$MC3_UsedUtility)
levels(MC3_CurrentLongDat$MC3_UsedUtility)

class(PriorStudyLong$MC1_PriorStrategyUtility)
PriorStudyLong$MC1_PriorStrategyUtility<-ordered(PriorStudyLong$MC1_PriorStrategyUtility, levels= c("Low", "Moderate", "High"))
PsGraph<- cbind.data.frame(PriorStudyLong$Strategies, PriorStudyLong$MC1_PriorStrategyUtility)
PsGraph<- na.omit(PsGraph)
colnames(PsGraph)<- c("Strategies", "MC1_PriorStrategyUtility")
ggplot(PsGraph, aes(Strategies, fill=MC1_PriorStrategyUtility))+ geom_bar() + theme_classic()+ scale_fill_viridis(option="D", discrete=TRUE)+facet_wrap(~MC1_PriorStrategyUtility)+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

Question 3: Does the use of effective strategies correlate with overall class performance?
We hypothesize that students with effective study strategies will earn higher grades in the class
We hypothesize that students with effective strategies will earn higher grades on exams than those who use ineffective strategies

```{r}
#cor(longdat$MC2_UsedUtility, longdat$Final.Grade)
#longdat$MC1_PriorStrategyUtility<- as.numeric(longdat$MC1_PriorStrategyUtility)
# need to use ordinal data correlation (spearman's) 
```

Question 4: Do students’ expected grades correlate with the actual grades that they earned? 
We hypothesize that students expected grades will correlate with their actual grades 
We hypothesize that there will be a greater difference (expected - actual) for those with poorer study strategies than those with more effective strategies 
```{r}
# Chi square Test ? 
#Mc2 actual vs expected
#Mc3 expected vs actual 
#for the paper, we may want to use reported numbers 
```

Question 5: Do students find reflecting on metacognition useful? 
We hypothesize that students will find reflecting on metacognition useful more often than chance
```{r}

```
