---
title: "Study Strategy Analyses"
author: "Rachel Rineman, Amanda Mae Woodward"
date: "9/29/2022"
output: html_document
---

```{r library}
library(readr)
library(expss)
library(tidyverse)
library(lme4)
```

```{r}
StudyDat  <- read_csv("Study Strategies Fixed.csv")
colnames(StudyDat)[1]<- "SubjectNumber"
colnames(StudyDat)[2]<- "ClassYear"
colnames(StudyDat)[4]<- "RaceEthnicity"
```

# Question 1: Do students in this class use effective study strategies? 
```{r priorSplit} 
PriorStudyDat<- separate(StudyDat,MC1_PriorStrategyDunlosky, into= c("PS1", "PS2", "PS3", "PS4", "PS5", "PS6"), sep= ",")
```

```{r switchtoLong }
PriorStudyLong<- pivot_longer(PriorStudyDat, cols= c("PS1", "PS2", "PS3", "PS4", "PS5", "PS6"), names_to= "orderMention", values_to= "Strategies")

PriorStudyLong$Strategies<- as.factor(PriorStudyLong$Strategies)
levels(PriorStudyLong$Strategies)<- c("Distributed", "Highlight", "Imagery", "Interleaved","Mnemonic","Mnemonic", "Reread", "SelfExplain", "Summary", "TestPractice","Distributed", "Highlight","Interleaved", "Mnemonic", "Reread","SelfExplain", "Summary", "TestPractice")

```

```{r Priorstudydatacleaning}
PriorStudyLong$MC1_PriorStrategyUtility<- ifelse(PriorStudyLong$Strategies == "Distributed", "High", PriorStudyLong$MC1_PriorStrategyUtility)
PriorStudyLong$MC1_PriorStrategyUtility<- ifelse(PriorStudyLong$Strategies == "TestPractice", "High", PriorStudyLong$MC1_PriorStrategyUtility)
PriorStudyLong$MC1_PriorStrategyUtility<- ifelse(PriorStudyLong$Strategies == "Interleaved", "High", PriorStudyLong$MC1_PriorStrategyUtility)
PriorStudyLong$MC1_PriorStrategyUtility<- ifelse(PriorStudyLong$Strategies == "Interrogation", "Moderate", PriorStudyLong$MC1_PriorStrategyUtility)
PriorStudyLong$MC1_PriorStrategyUtility<- ifelse(PriorStudyLong$Strategies == "SelfExplain", "Moderate", PriorStudyLong$MC1_PriorStrategyUtility)
PriorStudyLong$MC1_PriorStrategyUtility<- ifelse(PriorStudyLong$Strategies == "Reread", "Low", PriorStudyLong$MC1_PriorStrategyUtility)
PriorStudyLong$MC1_PriorStrategyUtility<- ifelse(PriorStudyLong$Strategies == "Highlight", "Low", PriorStudyLong$MC1_PriorStrategyUtility)
PriorStudyLong$MC1_PriorStrategyUtility<- ifelse(PriorStudyLong$Strategies == "Summary", "Low", PriorStudyLong$MC1_PriorStrategyUtility)
PriorStudyLong$MC1_PriorStrategyUtility<- ifelse(PriorStudyLong$Strategies == "Mnemonic", "Low", PriorStudyLong$MC1_PriorStrategyUtility)
PriorStudyLong$MC1_PriorStrategyUtility<- ifelse(PriorStudyLong$Strategies == "Imagery", "Low", PriorStudyLong$MC1_PriorStrategyUtility)
```

## Question 1: Do students in this class use effective study strategies?

### Do students use effective strategies more often than chance at the beginning of the course? 

#### Hypothesis: We hypothesize that students will not report using effective strategies at the beginning of the course

#### Descriptive Statistics About Study Strategies 
```{r descriptivePriorStrat}
cro(PriorStudyLong$Strategies)
library(viridis)
PriorStudyLong$MC1_PriorStrategyUtility<- as.factor(PriorStudyLong$MC1_PriorStrategyUtility)
levels(PriorStudyLong$MC1_PriorStrategyUtility)

class(PriorStudyLong$MC1_PriorStrategyUtility)
PriorStudyLong$MC1_PriorStrategyUtility<-ordered(PriorStudyLong$MC1_PriorStrategyUtility, levels= c("Low", "Moderate", "High"))
PsGraph<- cbind.data.frame(PriorStudyLong$Strategies, PriorStudyLong$MC1_PriorStrategyUtility)
PsGraph<- na.omit(PsGraph)
colnames(PsGraph)<- c("Strategies", "MC1_PriorStrategyUtility")
ggplot(PsGraph, aes(Strategies, fill=MC1_PriorStrategyUtility))+ geom_bar() + theme_classic()+ scale_fill_viridis(option="D", discrete=TRUE)+facet_wrap(~MC1_PriorStrategyUtility)+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

#### Descriptive Statistics About Study Strategy Utilities  
```{r descriptivePriorStratEffective}
cro(PriorStudyLong$MC1_PriorStrategyUtility)
ggplot(PsGraph, aes(MC1_PriorStrategyUtility))+geom_bar()+ theme_classic()
```
#of the 150 student responses (some students provided more than 1), 74 mentioned high utility 64 mentioned low, 10 mentioned moderate

#### Analysis: Generalized Linear Model Logistic Regression (Preliminary)
Dependent Variable: Utility (1= high utility; 0 = not high utility)
Independent Variable: Demographic Variables
```{r}
PriorStudyLong$Utility<- ifelse(PriorStudyLong$MC1_PriorStrategyUtility== "High", 1, 0)
# need to make column to indicate if student used *any* high utility strategy
colnames(PriorStudyDat)
PriorStudyDat<- as.data.frame(PriorStudyDat)
for(i in 5:10){
  PriorStudyDat[,i]<- as.factor(PriorStudyDat[,i])
}
class(PriorStudyDat[,5])
levels(PriorStudyDat[,5])[5]<- "Reread"
levels(PriorStudyDat[,6])<- c("Distributed","Highlight","Reread","SelfExplain","Summary","TestPractice")
levels(PriorStudyDat[,7])<- c("Distributed","Highlight","Imagery","Interleaved","Mnemonic","Reread","SelfExplain","Summary","TestPractice")
levels(PriorStudyDat[,8])<- c("Distributed","Highlight","Mnemonic","Reread","SelfExplain")
levels(PriorStudyDat[,9])<-c("Distributed","SelfExplain","TestPractice")
levels(PriorStudyDat[,10])<- c("SelfExplain","TestPractice")
PriorStudyDat$OverallUtility<-NA
```

```{r}
PriorStudyDat$PS1<-ifelse(is.na(PriorStudyDat$PS1)==TRUE, "none", levels(PriorStudyDat$PS1)[PriorStudyDat$PS1])
PriorStudyDat$PS2<-ifelse(is.na(PriorStudyDat$PS2)==TRUE, "none", levels(PriorStudyDat$PS2)[PriorStudyDat$PS2])
PriorStudyDat$PS3<-ifelse(is.na(PriorStudyDat$PS3)==TRUE, "none", levels(PriorStudyDat$PS3)[PriorStudyDat$PS3])
PriorStudyDat$PS4<-ifelse(is.na(PriorStudyDat$PS4)==TRUE, "none", levels(PriorStudyDat$PS4)[PriorStudyDat$PS4])
PriorStudyDat$PS5<-ifelse(is.na(PriorStudyDat$PS5)==TRUE, "none", levels(PriorStudyDat$PS5)[PriorStudyDat$PS5])
PriorStudyDat$PS6<-ifelse(is.na(PriorStudyDat$PS6)==TRUE, "none", levels(PriorStudyDat$PS6)[PriorStudyDat$PS6])

for (j in 5:10){
  for(i in 1:nrow(PriorStudyDat)){
    if(PriorStudyDat[i,j]== "TestPractice"){
      PriorStudyDat$OverallUtility[i]<-1
    }
    if(PriorStudyDat[i,j]=="Distributed"){
      PriorStudyDat$OverallUtility[i]<-1
    }
    if(PriorStudyDat[i,j]=="Interleaved"){
       PriorStudyDat$OverallUtility[i]<-1
    }
    else{print("whoops!")}
  }
}

PriorStudyDat$OverallUtility<- ifelse(is.na(PriorStudyDat$OverallUtility)==TRUE,0, PriorStudyDat$OverallUtility)
PriorStudyDat$ClassYear<- as.factor(PriorStudyDat$ClassYear)
levels(PriorStudyDat$ClassYear)[3]<- "First Year Student/Freshman" 
levels(PriorStudyDat$ClassYear)[4]<- "Second Year Student/Sophomore"

PriorStudyDat$Gender<-as.factor(PriorStudyDat$Gender)
levels(PriorStudyDat$Gender)[2]<- "Cisgender Woman" 
levels(PriorStudyDat$Gender)[4]<-"Do Not Wish to Disclose"
levels(PriorStudyDat$Gender)

PriorStudyDat$RaceEthnicity<- as.factor(PriorStudyDat$RaceEthnicity)
levels(PriorStudyDat$RaceEthnicity)

glm1<-glm(OverallUtility~ClassYear+RaceEthnicity+Gender, data=PriorStudyDat)
summary(glm1)
glm2<-glm(OverallUtility~ClassYear, data=PriorStudyDat)
summary(glm2)

```

### Analysis: Binomial Test 
```{r}
summary(as.factor(PriorStudyDat$OverallUtility))
binom.test(53, (53+13))
```
**Note:** Students are using highly effective strategies more often than chance before the course

```{r explorePrior}
PriorStudyDat$lowImpact<- NA

for (j in 5:10){
  for(i in 1:nrow(PriorStudyDat)){
    if(PriorStudyDat[i,j]== "Highlight"){
      PriorStudyDat$lowImpact[i]<-1
    }
    if(PriorStudyDat[i,j]=="Imagery"){
      PriorStudyDat$lowImpact[i]<-1
    }
    if(PriorStudyDat[i,j]=="Mnemonic"){
       PriorStudyDat$lowImpact[i]<-1
    }
     if(PriorStudyDat[i,j]=="Reread"){
       PriorStudyDat$lowImpact[i]<-1
     }
     if(PriorStudyDat[i,j]=="Summary"){
       PriorStudyDat$lowImpact[i]<-1
    }
    else{print("whoops!")}
  }
}

PriorStudyDat$lowImpact<- ifelse(is.na(PriorStudyDat$lowImpact)== TRUE, "none",PriorStudyDat$lowImpact)


summary(as.factor(PriorStudyDat$lowImpact))

binom.test(44,66)

# students are also statistically choosing low utility

PriorStudyDat$JustHighImpact<- ifelse(PriorStudyDat$OverallUtility==1 & PriorStudyDat$lowImpact !=1, 1, 0)
summary(as.factor(PriorStudyDat$JustHighImpact))
binom.test(14,66)

PriorStudyDat$JustLowImpact<- ifelse(PriorStudyDat$OverallUtility!=1 & PriorStudyDat$lowImpact==1, 1, 0)
summary(as.factor(PriorStudyDat$JustLowImpact))
binom.test(6,66)
```

#overall take away: more students than chance use high impact (and low impact strategies) at start of semester. Students seem to use a mix of both- not rely on one or the other 

Question 2: Will students use more effective strategies at the end of the course compared to at the beginning? 
We hypothesize that students will use more effective strategies as the semester progresses, compared to the beginning of the semester. 
### Data Clean 

#### MC 2
```{r}
colnames(StudyDat)
MC2StudyDat<- separate(StudyDat,MC2_UsedStrategyDunlosky, into= c("PS1", "PS2", "PS3", "PS4", "PS5", "PS6"), sep= ", ")

MC2StudyDat$PS1<-as.factor(MC2StudyDat$PS1)
levels(MC2StudyDat$PS1)[6]<- "Reread"

MC2StudyDat$PS2<-as.factor(MC2StudyDat$PS2)
levels(MC2StudyDat$PS2)<- c("Summary", "Distributed", "Distributed", "Highlight","Imagery","Interleaved","Interrogation","Mnemonic","Reread","Reread","SelfExplain", "Summary", "TestPractice")
MC2StudyDat$PS3<-as.factor(MC2StudyDat$PS3)
levels(MC2StudyDat$PS3)[4]<- "Reread"
MC2StudyDat$PS4<-as.factor(MC2StudyDat$PS4)
levels(MC2StudyDat$PS4)[2]<-"Reread" 
MC2StudyDat$PS5<-as.factor(MC2StudyDat$PS5)
levels(MC2StudyDat$PS5)
MC2StudyDat$PS6<-as.factor(MC2StudyDat$PS6)
levels(MC2StudyDat$PS6)
```

```{r}
MC2_CurrentLongDat<- pivot_longer(MC2StudyDat,cols= c("PS1", "PS2", "PS3", "PS4", "PS5", "PS6"), names_to= "orderMention", values_to= "Strategies")
```

#Longdat helps with graphs
```{r Priorstudydatacleaning}
colnames(MC2_CurrentLongDat)
MC2_CurrentLongDat$MC2_UsedUtility<- ifelse(MC2_CurrentLongDat$Strategies == "Distributed", "High", MC2_CurrentLongDat$MC2_UsedUtility)
MC2_CurrentLongDat$MC2_UsedUtility<- ifelse(MC2_CurrentLongDat$Strategies == "TestPractice", "High", MC2_CurrentLongDat$MC2_UsedUtility)
MC2_CurrentLongDat$MC2_UsedUtility<- ifelse(MC2_CurrentLongDat$Strategies == "Interleaved", "High", MC2_CurrentLongDat$MC2_UsedUtility)
MC2_CurrentLongDat$MC2_UsedUtility<- ifelse(MC2_CurrentLongDat$Strategies == "Interrogation", "Moderate", MC2_CurrentLongDat$MC2_UsedUtility)
MC2_CurrentLongDat$MC2_UsedUtility<- ifelse(MC2_CurrentLongDat$Strategies == "SelfExplain", "Moderate", MC2_CurrentLongDat$MC2_UsedUtility)
MC2_CurrentLongDat$MC2_UsedUtility<- ifelse(MC2_CurrentLongDat$Strategies == "Reread", "Low", MC2_CurrentLongDat$MC2_UsedUtility)
MC2_CurrentLongDat$MC2_UsedUtility<- ifelse(MC2_CurrentLongDat$Strategies == "Highlight", "Low", MC2_CurrentLongDat$MC2_UsedUtility)
MC2_CurrentLongDat$MC2_UsedUtility<- ifelse(MC2_CurrentLongDat$Strategies == "Summary", "Low", MC2_CurrentLongDat$MC2_UsedUtility)
MC2_CurrentLongDat$MC2_UsedUtility<- ifelse(MC2_CurrentLongDat$Strategies == "Mnemonic", "Low", MC2_CurrentLongDat$MC2_UsedUtility)
MC2_CurrentLongDat$MC2_UsedUtility<- ifelse(MC2_CurrentLongDat$Strategies == "Imagery", "Low", MC2_CurrentLongDat$MC2_UsedUtility)
cbind.data.frame(MC2_CurrentLongDat$Strategies, MC2_CurrentLongDat$MC2_UsedUtility)
```

```{r}
colnames(MC2_CurrentLongDat)
MC2Subset<- cbind.data.frame(MC2_CurrentLongDat$SubjectNumber, MC2_CurrentLongDat$orderMention, MC2_CurrentLongDat$Strategies, MC2_CurrentLongDat$MC2_UsedUtility) 
colnames(MC2Subset)<- c("SubjectNumber", "Order", "Strategies", "MC2_UsedUtility")

MC2SubsetWide<- pivot_wider(MC2Subset, names_from = "Order", values_from= "Strategies")
```



### MC 3 
```{r}
StudyDat$MC3_UsedDunloskyStrategy
MC3StudyDat<- separate(StudyDat,MC3_UsedDunloskyStrategy, into= c("PS1", "PS2", "PS3", "PS4", "PS5"), sep= ", ")

MC3StudyDat$PS1<- as.factor(MC3StudyDat$PS1)
levels(MC3StudyDat$PS1)[7]<- "Reread" 
levels(MC3StudyDat$PS1)[11]<- "TestPractice" 
MC3StudyDat$PS2<- as.factor(MC3StudyDat$PS2)
levels(MC3StudyDat$PS2)<- c("Distributed", "Highlight", "Highlight", "Imagery", "Interrogation", "Mnemonic", "Reread", "SelfExplain", "Summary", "TestPractice")
MC3StudyDat$PS3<- as.factor(MC3StudyDat$PS3)
levels(MC3StudyDat$PS3)[3]<- "Reread"
MC3StudyDat$PS4<- as.factor(MC3StudyDat$PS4)
levels(MC3StudyDat$PS4)
MC3StudyDat$PS5<- as.factor(MC3StudyDat$PS5)
levels(MC3StudyDat$PS5)
```

Question 3: Does the use of effective strategies correlate with overall class performance?
We hypothesize that students with effective study strategies will earn higher grades in the class
We hypothesize that students with effective strategies will earn higher grades on exams than those who use ineffective strategies

```{r}
#cor(longdat$MC2_UsedUtility, longdat$Final.Grade)
#longdat$MC1_PriorStrategyUtility<- as.numeric(longdat$MC1_PriorStrategyUtility)
# need to use ordinal data correlation (spearman's) 
```

Question 4: Do students’ expected grades correlate with the actual grades that they earned? 
We hypothesize that students expected grades will correlate with their actual grades 
We hypothesize that there will be a greater difference (expected - actual) for those with poorer study strategies than those with more effective strategies 
```{r}
# Chi square Test ? 
#Mc2 actual vs expected
#Mc3 expected vs actual 
#for the paper, we may want to use reported numbers 
```

Question 5: Do students find reflecting on metacognition useful? 
We hypothesize that students will find reflecting on metacognition useful more often than chance
```{r}

```

